---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 10

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Piyush Khedkar
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**Association mining** can be applied to many data problems beyond the well-known example of **finding relationships between different products in customer shopping data**. In this homework assignment, we will explore **real data** from the banking sector and look for **patterns associated with the likelihood of responding positively to a direct marketing campaign and signing up for a term deposit with the bank (stored in the variable  y )**. <br>
You can find out more about the variables in this dataset here: https://archive.ics.uci.edu/ml/datasets/bank+marketing

## Part 1: Explore Data Set

A.	Read the contents of the following URL to a dataframe called **bank** <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv

**Hint**: Even though this is a .csv file, chances are R won t be able to read it in correctly using the read_csv() function. If you take a closer look at the contents of the URL file, you may notice each field is separated by a **semicolon** (;) rather than a comma. 

In situations like this, consider using either **read.csv()** or **read.table()**, with two additional parameters. *sep=";"* defines how the data is seperated (the default is a comma), and *header=TRUE* defines that there is a header line in the dataset. 

```{r}
bank <- read.csv("https://intro-datascience.s3.us-east-2.amazonaws.com/bank-full.csv",
               sep = ";", header = TRUE)
```

B. Make sure there are **41,188** rows and **21** columns in your **bank** df.

```{r}
str(bank)
# Yes there are 41,188 rows and 21 columns
```

C.	Next, we will focus on some key factor variables from the dataset, and convert a few numeric ones to factor variables. Execute the following command.  Write a comment describing how the conversion for each numeric variable works and what are the variables in the resulting dataframe.

```{r}
bank_new <- data.frame(job=as.factor(bank$job),
                     marital=as.factor(bank$marital),
                     housing_loan=as.factor(bank$housing),
                     young=as.factor(bank$age<median(bank$age)),
                     contacted_more_than_once=as.factor(bank$campaign>1),
                     contacted_before_this_campaign=as.factor(bank$previous<0),
                     success=as.factor(bank$y))
head(bank_new)
# as.factor() converts a column from numeric to factor.
# job - Types of jobs (chr to factor)
# marital - married, single or divorced or unknown (chr to factor)
# housing_loan - yes or no (chr to factor)
# young -  IF age is less than median age then True else False (numeric to factor)
# contacted_more_than_once -  If campaign more than 1 then True else False(num to factor)
# contacted_before_this_campaign - IF prev less than 0 then True alse False(num to factor)
# succcess - Yes or no (chr to factor)
```

D.	Count the number of successful term deposit sign-ups, using the table( ) command on the **success** variable.

```{r}
 table(bank_new$success)
# 4640 successful term deposit sign-ups
```

E.	Express the results of problem C as percentages by sending the results of the table( ) command into the prop.table( ) command.

```{r}
prop.table(table(bank_new$success))*100
# 11.265% successful term deposit sign-ups and 88.734% unsuccesssful
```

F.	Using the same techniques, show the percentages for the **marital** and **housing_loan** variables as well.

```{r}
prop.table(table(bank_new$marital))*100
prop.table(table(bank_new$housing_loan))*100
# Percentages given below
```

## Part 2: Coerce the data frame into transactions

A.	Install and library two packages: **arules** and **arulesViz**.

```{r}
library(arules)
library(arulesViz)
```

B.	Coerce the **bank_new** dataframe into a **sparse transactions matrix** called **bankX**.

```{r}
bankX <-as(bank_new, 'transactions')
```

C.	Use the itemFrequency( ) and itemFrequencyPlot( ) commands to explore the contents of **bankX**. What do you see?

```{r}
itemFrequency(bankX)*100
# Shows frequency of items in percentage
itemFrequencyPlot(bankX)
# Barplot of frequencoes of factors between 0 to 1
# Admin jobs are most, unkown least
# More are married, more have housing loan, more people are not young
```

D.	This is a fairly large dataset, so we will explore only the first 10 observations in the **bankX** transaction matrix: 

```{r}
inspect(bankX[1:10])
# It shows details of first 10 people
```

E. Explain the difference between **bank_new** and **bankX** in a block comment:

```{r}
# bank_new is data frame of categorial data while bankX is sparse transaction 
# matrix. bank_new has 7 columns while sparse has 26 dimensions for example, in
# bank_new there's a column marital, but in bankX there will be 4 parts one
# each for  married, unmarried, divorced, unknown with 1 or 0 for T or F.
```

## Part 3: Use arules to discover patterns

**Support** is the proportion of times that a particular set of items occurs relative to the whole dataset. <br>
**Confidence** is proportion of times that the consequent occurs when the antecedent is present. <br>

A.	Use **apriori** to generate a set of rules with support over 0.005 and confidence over 0.3, and trying to predict who successfully signed up for a term deposit. <br>
**Hint:** You need to define the **right-hand side rule (rhs)**.

```{r}
ruleset <- apriori(bankX, 
                   parameter=list(supp=0.005, conf=0.3), 
                   control=list(verbose=F), 
                   appearance=list(default="lhs", rhs=("success=yes")))
```

B.	Use inspect() to review of the **ruleset**. 

```{r}
inspect(ruleset)
```

C.	Use the output of inspect( ) or inspectDT( ) and describe **any 2 rules** the algorithm found.  

```{r}
# Job is student has high support 0.00667 and confidence 0.314, so with success
# student has been occurred frequently.

# Job is student and when marital status is single it has high support and highest 
# confidence of 0.32, single and student occured frequently with success.
```

D. Generate a partition tree from the dataframe (not the transactions)

```{r}
library(rpart)
library(rpart.plot)

tree1<-rpart(success ~ ., data = bank_new)
rpart.plot(tree1)

tree2<-rpart(y ~ ., data = bank)
rpart.plot(tree2)
```

View the model (as a tree), and then explain with a block comment if this tree is helpful

```{r}
# For bank_new there is only one node which says that most rows in success are F
# For bank, nr.employed and duration columns have sub nodes but it does not tell
# us much about y column so both the trees are not very helpful.
```
 




