---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 6


##### Copyright Jeffrey Stanton 2022, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Piyush Khedkar
```


### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

**This module:** **Data visualization** is important because many people can make sense of data more easily when it is presented in graphic form. As a data scientist, you will have to present complex data to decision makers in a form that makes the data interpretable for them. From your experience with Excel and other tools, you know that there are a variety of **common data visualizations** (e.g., pie charts). How many of them can you name? <br>

The most powerful tool for data visualization in R is called **ggplot2**. Written by computer/data scientist **Hadley Wickham**, this ** graphics grammar ** tool builds visualizations in layers. This method provides immense flexibility, but takes a bit of practice to master.

## Step 1: Make a copy of the data

A.	Read the **New York State COVID Testing** dataset we used in HW 3 & 4 from this URL: https://data-science-intro.s3.us-east-2.amazonaws.com/NYS_COVID_Testing_.csv <br>
into a new dataframe called **df**. 


```{r}
library(tidyverse)
df = read.csv("https://data-science-intro.s3.us-east-2.amazonaws.com/NYS_COVID_Testing_.csv")
head(df)
```

B.	Your dataframe, **df**, contains a so-called **multivariate time series**: a sequence of measurements on COVID tests and results captured repeatedly over time (March 2020 - January 2022). Familiarize yourself with the nature of the time variable **TestDate**.
<br>How often were these measurements taken (in other words, at what frequency were the variables measured)? Put your answer in a comment. 

```{r}
table(df$AgeGroup) #There are 11 age categories
# summary(df) #There are 7838 times te measurements were taken, some days where
# multiple measurements taken on same day for different age caetegories. 
```

C. What is the data type of **TestDate**? Explain in a comment.

```{r}
class(df$TestDate)
# The data type of TestDate is character
```

D. To properly display the **TestDate** values as dates in our plots, we need to convert **TestDate** to date format with the **as.Date()** function. Run the code below and check the data type of the variable again to make sure it is not coded as text anymore:

```{r}
df$TestDate<-as.Date(df$TestDate, format = "%m/%d/%Y")
class(df$TestDate)
```

```{r}
# The data type of TestDate is now Date
```

## Step 2: Clean up the NAs and create subsets

A.	It is always good practice, when you first start working with a dataset, to explore it for missing values. Check the **TotalTests** and **PositiveCases** for missing values. Are there any? What does empty output suggest about the number of missing observations?

Hint: use *is.na()* 

```{r}
sum(is.na(df$TotalTests))
# 22 missing values in Total Tests
sum(is.na(df$PositiveCases))
# 11 missing values in Positive Cases
```

B. There is an R package called **imputeTS** specifically designed to repair missing values in time series data. We will use this instead of the simpler way, **mean substitution**, because it tends to be more accurate.
<br>The **na_interpolation()** function in this package takes advantage of a unique characteristic of time series data: neighboring points in time can be used to  guess  about a missing value in between. <br>Use this function on each of the two numeric variables in **df** and don't forget to **update** them by overwriting them with the output of the **na_interpolation()** function. 

```{r}
# install.packages("imputeTS", dependencies = TRUE)
# install.packages("tseries", dependencies = TRUE)
library(imputeTS)
length(na_interpolation(df$TotalTests))
length(na_interpolation(df$PositiveCases))
df$TotalTests <- na_interpolation(df$TotalTests)
df$PositiveCases <- na_interpolation(df$PositiveCases)
```

C. Run the code from A to make sure there is no more missing data:

```{r}
sum(is.na(df$TotalTests))
sum(is.na(df$PositiveCases))
```

D. As we've done before, let's create a new variable which is the ratio of **PositiveCases** to **TotalTests** - save it as an additional variable in **df** called **PositivityRate**:

```{r}
df$PositivityRate <- df$PositiveCases/df$TotalTests
length(df$PositivityRate)
```

E. Create a subset of **df** containing **only the records for children**. Save it in a new dataframe called **dfChildren**. Make sure this new df has **2,010 observations and 8 variables**.

```{r}
dfChildren <- df[df$AgeCategory == 'children',]
summary(dfChildren)
```

F. Create a subset of **df** containing only the records for **young adults**. Save it in a new dataframe called **dfYA**. 

```{r}
dfYA <- df[df$AgeCategory == 'young_adults',]
```

G. Using the same logic, create 2 more subsets of **df**: one containing only the records for **middle-aged adults** (call it **dfMA**), and another one with only the data of **older adults** - **dfOA**. After this step, you should have a total of 4 subsets: 
- dfChildren
- dfYA
- dfMA
- dfOA

```{r}
dfMA <- df[df$AgeCategory == 'middle-aged_adults',]
dfOA <- df[df$AgeCategory == 'older_adults',]
```

## Step 3: Use ggplot to explore the distribution of each variable
**Don t forget to install and library the ggplot2 package.** Then: <br>

A.	Create a histogram for **PositiveCases** in the **dfOA** dataframe (using **ggplot**). Be sure to add a title and briefly describe what the histogram means in a comment.

```{r}
library(ggplot2)
ggplot(dfOA, aes(x=PositiveCases)) + geom_histogram() + ggtitle("Frequency of Positive Cases in Older Adults")
# PositiveCases highest between 0 to 625
```

B.	Create histograms (using **ggplot**) of the **PositiveCases** variable in each of the other three subsets from Step 2G. <br>
For each histogram, comment on its shape - what information can we glean from it?

```{r}
ggplot(dfChildren, aes(x=PositiveCases)) + geom_histogram() + ggtitle("Frequency of Positive Cases in Children")
# PositiveCases highest between 0 to 1250
```

```{r}
ggplot(dfMA, aes(x=PositiveCases)) + geom_histogram() + ggtitle("Frequency of Positive Cases in Middle Aged Adults")
# PositiveCases highest between 0 to 1250
```

```{r}
ggplot(dfYA, aes(x=PositiveCases)) + geom_histogram() + ggtitle("Frequency of Positive Cases in Young Adults")
# PositiveCases highest between 0 to 1250
```

## Step 4: Explore how the data changes over time
A.	These data were collected in a period of almost 2 years. You can thus observe changes over time with the help of a line chart. Let's focus on the **dfOA** subset first: <br>
Create a **line chart**, with **TestDate** on the X-axis and **PositivityRate** on the Y-axis.

```{r}
ggplot(dfOA, aes(x = TestDate, y = PositivityRate)) + geom_line() + ggtitle("Positivty Rate over time")
# PositivityRate spiked after 2020-04,05 and then dipped to spike a bit in 2021-01
# and then dipped to spike up again in 2022-01
```

B.	Next, create similar graphs for each of the other three subsets. Change the **color** of the line plots (any color you want).

```{r}
ggplot(dfChildren, aes(x = TestDate, y = PositivityRate)) + geom_line(color = "orange") + ggtitle("Positivty Rate over time")
# PositivityRate spiked after 2020-04,05,06 and then dipped to spike a bit in 2021-01
# and then dipped in 2021-07 to spike up again in 2022-01
```

```{r}
ggplot(dfMA, aes(x = TestDate, y = PositivityRate)) + geom_line(color = "purple") + ggtitle("Positivty Rate over time")
# PositivityRate spiked after 2020-04 and then 06 and then dipped to spike a bit in 2021-01
# and then dipped in 2021-07 to spike up again in 2022-01
```

```{r}
ggplot(dfYA, aes(x = TestDate, y = PositivityRate)) + geom_line(color = "blue") + ggtitle("Positivty Rate over time")
# PositivityRate spiked after 2020-04 and then dipped to spike a bit in 2021-01
# and then dipped in 2021-07 to spike up again in 2022-01
```

```{r}

```

C.	In a comment, talk about the insights you got from the line charts you created - can you spot any trends within and between the line charts?

```{r}
# Though the Positivty Rate touched and crossed 1 in Children, YA it remained
# on 0.4, 0.6 for MA, OA. For all they spiked in the biginning of 2020, then
# dipped to spike a bit in 2021-01 and then dipped in 2021-07 to spike up 
# again in 2022-01
```

D.	Finally, using the **dfOA** subset, create a **scatter plot**, showing **TotalTests** on the x axis, **PositiveCases** on the y axis, and having the **color and size** of the point represent **Year**.

```{r}
ggplot(dfOA, aes(x=TotalTests,y=PositiveCases)) + geom_point(aes(color = Year, size=Year))
```

E. Create a similar scatter plot for the **dfYA** subset. 

```{r}
ggplot(dfYA, aes(x=TotalTests,y=PositiveCases)) + geom_point(aes(color = Year,size=Year))
```

F.	Create two barcharts (**using ggplot**) that you think would be interesting by exploring any attribute in two of the dataframes that you have already created via a barchart.

```{r}
ggplot(data=dfChildren, aes(x=AgeGroup, y=PositiveCases)) + geom_bar(stat="identity")+ ggtitle("Barchart for Children")

ggplot(data=dfOA, aes(x=AgeGroup, y=PositiveCases)) + geom_bar(stat="identity")+ ggtitle("Barchart for OA")
```

G.	Interpret these visualizations, what insight do they provide?

```{r}
# For Children, more PositiveCases for ages 5 to 19, then 1 to 4, then <1
# For OA  more PositiveCases for ages 55 to 64, then 65 to 74 and least for 85 +
# The reason is Children below 1 probably did not interact with a lot of people
# just like 85+ but 55 to 64, 5 to 19 would be socially active.
```


