---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 2

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Piyush Khedkar
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

### Reminders of things to practice from last week: 
Assignment arrow:  		<- <br>
The combine command: 	c( ) <br>
Descriptive statistics: 	mean( ) sum( ) max( ) <br>
Arithmetic operators:		+ - * /<br>
Boolean operators:		>    <    >=    <=   ==   !=  <br>

**This Week:** Explore the **quakes** dataset (which is included in R). 


<br>
<br>

**Step 0:** Prepare the data <br>
A. Copy the **quakes** dataset into a new dataframe (call it **myQuakes**), so that if you need to start over, you can do so easily (by copying **quakes** into **myQuakes** again). Summarize the variables in **myQuakes** using an R command. Also, explore the structure of the dataframe - using another R command. In a brief comment, describe what this dataset is about based on your analysis of the variables in it. 

```{r}
myQuakes <- quakes
summary(myQuakes)
# This dataset is about earthquakes, latitude, longitude of where they happened,
# their depth, magnitude and no of stations reporting. 
```

**Step 1:** Explore the earthquake magnitude variable called **mag**  <br>

A.	What is the average magnitude? Use mean() or summary():

```{r}
mean(myQuakes$mag)
```

B.	What is the magnitude of the smallest earthquake? Use min() or summary() and save the result in a variable called **minQuake**:

```{r}
minQuake <- min(myQuakes$mag)
minQuake
```

C.	What is the magnitude of the largest earthquake? Use max() or summary() and save the result in a variable called **maxQuake**:

```{r}
maxQuake <- max(myQuakes$mag)
maxQuake
```

D.	Output the **fourth row** of the dataframe:

```{r}
myQuakes[4,]
```

E. Create a new dataframe, with only the rows where the **magnitude is greater than 4**. How many rows are in that dataframe (report this number using code, do not count by looking at the output):

```{r}
mag_greater <- myQuakes[myQuakes$mag > 4,]
nrow(mag_greater)
```

F.	Using the **arrange()** function, create a **sorted dataframe** based on magnitude and store it in **quakeSorted1**. Explain how this function works.

```{r}
library("tidyverse")
quakeSorted1 <-arrange(myQuakes, myQuakes$mag)
head(quakeSorted1)
# Arrange() sorts rows of df by the given column, default is ascending
```

G. Now create a second sorted dataframe called **quakeSorted2** using the **order()** function. Explain in a comment how the function works.

```{r}
indexes <- order(myQuakes$mag)
quakeSorted2 <- myQuakes[indexes,]
#row.names(quakeSorted2) <- NULL
head(quakeSorted2)
# Order() returns indices of sorted values by given column(default is ascending)
# then these indices have to be used to save value in new df.
```

H. Print the top rows of each of the sorted dataframes using the **head()** function and explain in a comment what you see - did the two sorting functions produce a similar result?

```{r}
head(quakeSorted1)
head(quakeSorted2)
```

```{r}
# The values in each row are same, but the row names have been changed because
# indexes of myQuakes are used in quakeSorted1, they need to be reordered.
```

I.	What are the latitude and longitude of the quake reported by the largest number of stations?  

```{r}
max_stations <- which.max(myQuakes$stations)
myQuakes[max_stations, c("lat","long")]
```

J.	What are the latitude and longitude of the quake with the smallest magnitude?  

```{r}
min_stations <- which.min(myQuakes$stations)
myQuakes[min_stations, c("lat","long")]
```

**Step 3:** Using conditional if statements <br>

A.	 Test if **maxQuake** is greater than 6 (output  yes  or  no ) <br>
**Hint:** Try modifying the following code in R: 

```{r}
if  (maxQuake > 6) "yes" else "no"
```

B.	 Following the same logic, test if **minQuake** is less than 3 (output  yes  or  no ):

```{r}
if (minQuake < 3) "yes" else "no"
```

