---
output:
  pdf_document: default
  html_document: default
---

## Intro to Data Science - HW 1
##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Piyush Khedkar
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Define a variable:

```{r}
value <- 150
```

Print the content of this new variable:

```{r}
print(value)
```

Define the following vectors, which represent the **hospital capacity** (measured in number of hospital beds) and **number of colleges** in each of the five counties in Central New York (CNY)   **Cayuga**, **Cortland**, **Madison**, **Onondaga**, and **Oswego**, in this order (e.g. the first number in the **capacity** vector suggests there are 124 hospital beds in Cayuga county):

```{r}
capacity <- c(124, 70, 51, 1403, 86)
colleges <- c(2, 2, 3, 9, 2)
```

# Part 1: Calculating statistics using R  

A.	Show the number of observations in the **capacity** vector with the length() function: 

```{r}
no_of_hospital_county <- length(capacity)
no_of_hospital_county
```

B.	Show the number of observations in the **colleges** vector with the length() function:

```{r}
no_of_colleges <- length(colleges)
no_of_colleges
```

C.	Calculate the average number of colleges in CNY using the mean() function:

```{r}
mean(colleges)
```

D.	Calculate the total number of colleges in CNY using the sum() function:

```{r}
sum_colleges <- sum(colleges)
sum_colleges
```

E.	Calculate the average CNY hospital capacity using the mean() function:

```{r}
mean(capacity)
```

F.	Calculate the total hospital capacity in CNY using the sum() function:

```{r}
total_hospital_capacity <- sum(capacity)
total_hospital_capacity
```

G.	Calculate the average number of colleges in CNY again, this time using **the results from steps B & D**:

```{r}
average_no_of_colleges <- sum_colleges/no_of_colleges
average_no_of_colleges
```

H.	Calculate the average CNY hospital capacity again, this time using **the results from steps A & F**:

```{r}
average_hospital_capacity <- total_hospital_capacity/no_of_hospital_county
average_hospital_capacity
```

# Part 2: Using the max/min and range functions in R

I.	How many colleges does the county with most colleges have? Hint: Use the max() function:

```{r}
max(colleges)
```

J.	What is the hospital capacity of the county with the smallest number of hospital beds in CNY? **Hint:** Use the min() function:

```{r}
min(capacity)
```

K.	Display the number of beds in the county with the highest hospital capacity and the county with the lowest hospital capacity together, using a single command. **Hint:** Use the range() function:

```{r}
range(capacity)
```

# Part 3: Vector Math

L.	Create a new vector called **lowerCap**, which is the current hospital capacity of each county **- 10** (each county has 10 less hospital beds):

```{r}
lowerCap <- capacity - 10
lowerCap
```

M.	 Calculate the average of **lowerCap**:

```{r}
mean(lowerCap)
```

N. Create another vector called **megaCap**, which is the original hospital capacity in each county (stored in **capacity**) plus **value** (the variable we declared in the beginning of this assignment). 

```{r}
megaCap <- capacity + value
megaCap
```

O. Print the content of **megaCap**:

```{r}
print(megaCap)
```

P.	In a variable called **bigCap**, store all the hospital numbers from the original **capacity** vector which are **greater than 80** (using **subsetting** in R):

```{r}
bigCap <- capacity[capacity > 80]
bigCap
```

Q.	Print the length of **bigCap** and explain in a comment how many counties have more than 80 hospital beds:

```{r}
print(length(bigCap))
# Since bigCap vector has all the counties with capacity greater than 80 beds, the length
#of bigCap will give the number of counties that have more than 80 hospital bed which is 3.
```

