---
output:
  pdf_document: default
  html_document: default
---

# Intro to Data Science - HW 5

##### Copyright 2022, Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva

```{r}
# Enter your name here: Piyush Khedkar
```

### Attribution statement: (choose only one and delete the rest)

```{r}
# 1. I did this homework by myself, with help from the book and the professor.
```

Reminders of things to practice from previous weeks: <br>
Descriptive statistics: 	mean( ) max( ) min( )<br>
Coerce to numeric:		as.numeric( )<br>

## Part 1: Use the Starter Code 

Below, I have provided a starter file to help you. <br>

Each of these lines of code **must be commented** (the comment must that explains what is going on, so that I know you understand the code and results).

```{r}
#install.packages('RCurl')
#install.packages('jsonlite')
library(RCurl)
library(jsonlite)
# Now we can use RCurl, jsonlite libraries
dataset <- getURL("https://intro-datascience.s3.us-east-2.amazonaws.com/role.json")
readlines <- jsonlite::fromJSON(dataset)
# Convert from JSON to R object list
df <- readlines$objects$person
#Selected only values from objects and person and stored into dataframe
```

A.	Explore the **df** dataframe (e.g., using head() or whatever you think is best).

```{r}
str(df)
```

B.	Explain the dataset <br>
o	What is the dataset about? <br>
o	How many rows are there and what does a row represent?<br>
o	How many columns and what does each column represent?<br>

```{r}
# The dataset has information of senators
# There are 100 rows of data
# there are 17 columns, with each senator's birth date, id, name, gender, link to 
# profile, their twitter and youetube id
```

## Part 2: Investigate the resulting dataframe 

A.	Describe what you see when you run the **table()** function on the **gender** variable.

```{r}
table(df$gender)
# It builds the contingency table of objects present on column, gives count of
# frequency of occurence in %
help(table)
```

A1.	Generate the count of number of females and number of males, using the tidyverse **group_by()**, **summarise()** and **n()** functions.

```{r}
library(tidyverse)
df%>%group_by(gender)%>%summarise(gender_count = n())
```

B. How many senators are women?

```{r}
# 24 of them are senators as seen above
```

C.	How many senators don't have a YouTube account? <br> **Hint:** You can use the **is.na** function to locate the rows for which the YouTube account is missing and then wrap it in the **nrow()** or **sum** function to count the number of missing instances.

```{r}
nrow(df[is.na(df$youtubeid),])
# 27 senators do not have a Youtube id
```

D. Using the approach in C, i.e.using the **is.na()** function, show how many senators ***do*** have a YouTube account. <br>
**Hint:** You can reverse the **is.na()** function by placing a **!** in front of it - **!is.na( )**.  

```{r}
nrow(df[!is.na(df$youtubeid),])
# 73 of them have youtube account
```

E.	How many women senators have a YouTube account? 

```{r}
df%>%filter(gender=='female' & !is.na(youtubeid))%>%summarise(no_women_senators=n())
# 16 women senators have a youtube id
```

F.	Create a new dataframe called **youtubeWomen** that only includes women senators who have a YouTube account.   

```{r}
youtubeWomen <- df%>%filter(gender=='female' & !is.na(youtubeid))
head(youtubeWomen)
```

G.	What does running this line of code do? Explain in a comment: 

```{r}
youtubeWomen$year <- substr(youtubeWomen$birthday,1,4)
youtubeWomen$year
#It takes a substring of birthdat for 1st four characters, i.e only years
```

H.	Use this new variable to calculate the mean **birthyear** in **youtubeWomen**. <br>
**Hint:** You may need to convert it to numeric first using the **as.numeric()** function.

```{r}
mean(as.numeric(youtubeWomen$year))
```

I.	Make a histogram of the **birthyears** of senators in **youtubeWomen**. Add a comment describing the shape of the distribution.

```{r}
hist(as.numeric(youtubeWomen$year))
# It looks like a normal bell shape, with most of female senators are born 
# between 1950 and 1960, followed by 1940-50
```

J. Create a dataframe called **youtubeMen** which only includes male senators with a youTube account. Repeat steps G & H for this dataframe and create a histogram of the birthyears in it. Compare the shape and properties of this histogram to the one in H.

```{r}
youtubeMen <- df%>%filter(gender=='male' & !is.na(youtubeid))
youtubeMen$year <- substr(youtubeMen$birthday,1,4)
hist(as.numeric(youtubeMen$year))
# Male senators are born between 1950-55 followed by 1945-50 is almost a 
# normal curve with some anomalies
```

K. Take a look at [this article](https://www.theguardian.com/us-news/ng-interactive/2018/nov/15/new-congress-us-house-of-representatives-senate) - explore its interactive features and focus specifically on the section on **gender**. Relating what you learned from the article back to our Senate data, who might feel left out and/or unrepresented based on the current gender composition of the Senate? Explain in a brief comment.

```{r}
# Congress has 409 male and 126 females that is 76.4% male and 23.6% are women
# senators similar to our data. There are no transgender/non-binary senators 
# either in Congress, Democrats or Republicans while 0.6% of Americans are 
# transgender. They might feel left out.
# Also women senators are less than 50% in Congress, Democrats, Republicans
# despite the total population, they might be under represented.

```

